---
title: "Biodiversity Database and Biodiversity Indexing"
author:   
  - Deependra Dhakal 
institute:   
  - College of Natural Resource Management, Tikapur, Kailali
  - Agriculture and Forestry University 
output:   
  beamer_presentation:  
    incremental: false  
    theme: "Frankfurt"  
    colortheme: "beaver"
    toc: true   
    slide_level: 2
    keep_tex: true
    includes:
      in_header: 09-biodiversity_database_indexing_beamer_header.tex
classoption: "aspectratio=169"
header-includes: 
- \AtBeginSubsection{}
bibliography: [../bibliography/bibliographies.bib]
---

```{r setup, include=FALSE}
library(knitr)
require(tidyverse)
set.seed(453)
# invalidate cache when the package version changes
knitr::opts_chunk$set(tidy = FALSE, echo = FALSE, 
                  message = FALSE, warning = FALSE,
                  out.width = "45%")
options(knitr.table.format = "latex")
options(knitr.kable.NA = "", digits = 2)
options(kableExtra.latex.load_packages = FALSE)
```

# Biodiversity Database

## 

# Biodiversity Indices

## Meaning

\footnotesize

A diversity **index** is a quantitative measure that reflects **how many different types** (such as species) there are in a dataset (a community), and that can simultaneously take into account the phylogenetic relations among the individuals distributed among those types, such as richness, divergence or evenness. These indices are statistical representations of biodiversity in different aspects (richness, evenness, and dominance). 

When diversity indices are used in ecology, the types of interest are usually species, but they can also be other categories, such as genera, families, functional types, or haplotypes. The entities of interest are usually individual plants or animals, and the **measure of abundance** can be, for example, number of individuals, biomass or coverage.

The most commonly used diversity indices are simple transformations of the **effective number of types** (also known as '**true diversity**').

## Effective number of species

\footnotesize

**True diversity**, or the **effective number of types**, refers to the number of equally abundant types needed for the average proportional abundance of the types to equal that observed in the dataset of interest (where all types may not be equally abundant). The true diversity in a dataset is calculated by first taking the weighted generalized mean $M_{q-1}$ of the proportional abundances of the types in the dataset, and then taking the reciprocal of this. The equation is:

$$
{}^{q}\!D={1 \over M_{q-1}}=
{1 \over {\sqrt[{q-1}]{\sum_{i=1}^{R}p_{i}p_{i}^{q-1}}}}=\left({\sum_{i=1}^{R}p_{i}^{q}}\right)^{1/(1-q)}
$$

The denominator $M_{q-1}$ equals the average proportional abundance of the types in the dataset as calculated with the weighted generalized mean with exponent $q-1$. In the equation, R is richness (the total number of types in the dataset), and the proportional abundance of the $i$th type is $p_i$. The proportional abundances themselves are used as the nominal weights. The numbers ${}^{q}D$ are called Hill numbers of order $q$ or effective number of species.

## Shannon entropy

When q = 1, the above equation is undefined. However, the mathematical limit as q approaches 1 is well defined and the corresponding diversity is calculated with the following equation:

$$
{}^{1}\!D={1 \over {\prod _{i=1}^{R}p_{i}^{p_{i}}}}=\exp \left(-\sum _{i=1}^{R}p_{i}\ln(p_{i})\right)
$$

which is the exponential of the **Shannon entropy** calculated with natural logarithms.

## General Equation

$$
{}^{q}\!D=\left({\sum _{i=1}^{R}p_{i}^{q}}\right)^{1/(1-q)}
$$

The term inside the parentheses is called the basic sum. Some popular diversity indices correspond to the basic sum as calculated with different values of q.

## Sensitivity of diversity value to rare versus abundant species

\footnotesize

It defines the sensitivity of the true diversity to rare vs. abundant species by modifying how the weighted mean of the species' proportional abundances is calculated. With some values of the parameter $q$, the value of the generalized mean $M_{q-1}$ assumes familiar kinds of weighted means as special cases. In particular,

- $q = 0$ corresponds to the weighted harmonic mean,
- $q = 1$ to the weighted geometric mean, and
- $q = 2$ to the weighted arithmetic mean.
- As $q$ approaches infinity, the weighted generalized mean with exponent $q-1$ approaches the maximum $p_i$ value, which is the proportional abundance of the most abundant species in the dataset.

Generally, increasing the value of $q$ increases the effective weight given to the most abundant species. This leads to obtaining a larger $M_{q-1}$ value and a smaller true diversity (${}^q D$) value with increasing $q$.

## Richness

- Richness $R$ simply quantifies how many different types the dataset of interest contains. For example, species richness (usually noted $S$) of a dataset is the number of species in the corresponding species list.
- Richness is a simple measure, so it has been a popular diversity index in ecology, where abundance data are often not available for the datasets of interest.
- Because richness does not take the abundance of the types into account, it is not the same thing as diversity, which does take abundance into account.
- If true diversity is calculated with $q = 0$, the effective number of types (${}^0 D$) equals the actual number of types, which is identical to Richness ($R$).

## Shannon Index

- The Shannon index = Shannon's diversity index = Shannon–Wiener index $\neq$  Shannon–Weaver index (erroneous).
- The measure was originally proposed by Claude Shannon in 1948 to quantify the entropy. 

It is most often calculated as follows:

$$
H'=-\sum _{i=1}^{R}p_{i}\ln p_{i}
$$
$p_i$ is often the proportion of individuals belonging to the $i$th species in the dataset of interest. Then the Shannon entropy quantifies the uncertainty in predicting the species identity of an individual that is taken at random from the dataset.

##

The Shannon index ($H'$) is related to the weighted geometric mean of the proportional abundances of the types. Specifically, it equals the logarithm of true diversity as calculated with $q = 1$:

$$
H'=-\sum _{i=1}^{R}p_{i}\ln p_{i}=-\sum _{i=1}^{R}\ln p_{i}^{p_{i}}
$$
This can also be written

$$
H'=-(\ln p_{1}^{p_{1}}+\ln p_{2}^{p_{2}}+\ln p_{3}^{p_{3}}+\cdots +\ln p_{R}^{p_{R}})
$$

which equals,

##

\footnotesize

$$
H'=-\ln p_{1}^{p_{1}}p_{2}^{p_{2}}p_{3}^{p_{3}}\cdots p_{R}^{p_{R}}=\ln \left({1 \over p_{1}^{p_{1}}p_{2}^{p_{2}}p_{3}^{p_{3}}\cdots p_{R}^{p_{R}}}\right)=\ln \left({1 \over {\prod _{i=1}^{R}p_{i}^{p_{i}}}}\right)
$$

Since the sum of the pi values equals unity by definition, the denominator equals the weighted geometric mean of the $p_i$ values, with the $p_i$ values themselves being used as the weights (exponents in the equation). The term within the parentheses hence equals true diversity ${}^1 D$, and $H'$ equals $ln({}^1 D)$.

When all types in the dataset of interest are equally common, all $p_i$ values equal $\frac{1}{R}$, and the Shannon index hence takes the value $ln(R)$. The more unequal the abundances of the types, the larger the weighted geometric mean of the $p_i$ values, and the smaller the corresponding Shannon entropy. If practically all abundance is concentrated to one type, and the other types are very rare (even if there are many of them), Shannon entropy approaches zero. When there is only one type in the dataset, Shannon entropy exactly equals zero (there is no uncertainty in predicting the type of the next randomly chosen entity). 

# Numerical problems

## Problem

1. Calculate Shannon Weaner Index, Simpson Index and Evenness from following two areas and interpret your result. Also discuss species richness and evenness from following information.

```{r insect-area-comparison}
tribble(~"Insect", ~"Area 1", ~"Area 2", 
        "Wasp", 112, 424,
        "Butterfly", 88, 76, 
        "Grasshopper", 143, 54, 
        "Bettle", 112, 60, 
        "Bee", 100, 40, 
        "Hoverfly", 145, 46) %>% 
  t() %>% 
  magrittr::set_colnames(.[1,]) %>% 
  as_tibble() %>% 
  slice(-1) %>% 
  knitr::kable(booktabs = TRUE, caption = "Data of insect counts from two areas.") %>% 
  kableExtra::kable_styling(font_size = 8)
```

## Solution

\begin{columns}[T,onlytextwidth]
  \column{0.5\textwidth}

Here, 

We have Shannon-Weiner Index,

$$ 
H = \sum_{i = 1}^S{p_i.\ln p_i}
$$

Where, $S$ is the species richness (number of distinct species types), $p_i$ is the proportional abundance, and $\ln p_i$ the natural log of the proportional abundance.

  \column{0.5\textwidth}

```{r insect-area-index-calculation}
tribble(~"Insect", ~"Area 1", ~"Area 2", 
        "Wasp", 112, 424,
        "Butterfly", 88, 76, 
        "Grasshopper", 143, 54, 
        "Bettle", 112, 60, 
        "Bee", 100, 40, 
        "Hoverfly", 145, 46) %>% 
  mutate_at(c("Area 1", "Area 2"), list(pi = ~./(sum(.)), lnpi = ~log(./(sum(.))))) %>% 
  summarise(A1_d_shannon_weiner = -sum(`Area 1_lnpi` * `Area 1_pi`), 
            A1_d_simpson = -sum(`Area 1_pi`^2), 
            A1_d_difference_simpson = 1-A1_d_simpson,
            A1_d_richness = n(),
            A1_d_evenness = A1_d_shannon_weiner/log(A1_d_richness),
            A2_d_shannon_weiner = -sum(`Area 2_lnpi` * `Area 2_pi`), 
            A2_d_simpson = -sum(`Area 2_pi`^2), 
            A2_d_difference_simpson = 1-A2_d_simpson, 
            A2_d_richness = n(),
            A2_d_evenness = A2_d_shannon_weiner/log(A2_d_richness)
            ) %>% 
  pivot_longer(cols = A1_d_shannon_weiner:A2_d_evenness, 
               names_to = c("Area", "Index"), 
               names_pattern = "A(\\d)_d_(.*)"
               ) %>% 
  pivot_wider(names_from = Area, names_prefix = "Area ") %>% 
  knitr::kable(booktabs = TRUE, caption = "Computed diversity indices for insect count data.") %>% 
  kableExtra::kable_styling(font_size = 8, position = "center")
```

\end{columns}
